\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-01-09 archive.tex
more progress on archive

\section{Wake/Sleep Program Induction}\label{overviewSection}

@@

Viewed as an inference problem, our goal is to start with a collection
of tasks, written $X$, and infer both a program for each task, as well
as a distribution over programs encoded by a DSL, written
$\mathcal{D}$.  We equip $\mathcal{D}$ with a learned weight vector
$\theta$, and together $(\mathcal{D},\theta)$ define a generative
model over programs (see Appendix~\ref{sampleProgram}).  We perform
maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given
$X$.  Writing $J$ for the joint probability of $(\mathcal{D},\theta)$
and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]&\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta &\qquad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
where $\probability[x|p]$ scores the likelihood of a task
$x\in X$ given a program $p$.\footnote{For example, for list
  processing, the likelihood is 1 if the program predicts the observed
  outputs on the observed inputs, and 0 otherwise; when learning a
  generative model or probabilistic program, the likelihood is the
  probability of the program sampling the observation.}
The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:

\begin{definition}
A \textbf{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
\end{definition}
\noindent Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}
Our wake and sleep cycles correspond to alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$  (\textbf{Wake})
and $(\mathcal{D},\theta)$ (\textbf{Consolidation}):
\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
is large (i.e., programs with high posterior probability).
%Here we use a simple and generic enumerative program synthesis algorithm.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\noindent\textbf{Sleep (Dreaming): tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$Q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
into the frontiers will most increase $\lowerBound$.

@@

During waking we enumerate programs from the DSL in decreasing order
of their probability according to the recognition model, and then
check if a program $p$ assigns positive probability to a task
($\probability[x|p] > 0$); if so, we incorporate $p$ into the frontier
$\mathcal{F}_x$ (Figure~\ref{neuralPipeline}).  We represent programs
as polymorphicly-typed $\lambda$-calculus expressions, a
representation closely resembling Lisp and functional languages like
Haskell and OCaml, including variables, conditionals, higher-order
recursive functions, and the ability to create new functions.
We `batch' the tasks by randomly shuffling the training set
and playing small minibatches
of tasks to the agent during each wake cycle.

@@

Why enumerate, when the program synthesis community has invented many
sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
We have two reasons:
(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
the space of  programs.
\begin{figure}
  \begin{tikzpicture}[every node/.style={rounded corners,thick}]
    \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
    \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
      e.g., RNN or CNN};
%    \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
    \node[draw,align = center](s) at ([xshift=2cm]e.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
    \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Frontier\\$\mathcal{F}_x$};

    \draw[->] (x.east) -- (e.west);
    \draw[->] (e.east) -- (s.west);
%    \draw[->] (m.east) -- node(p)[fill=white,align=center,midway,inner sep=0,outer sep=0,rotate=90]{\small $Q_{ijk}(x)$}(s.west);
    \draw[->] (s.east) -- (f.west);
    \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway]{\small likelihood $\probability[x|p]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
    \node[align=center] at ($(0,-1) + (e.center)$) {$\underbrace{\hspace{4cm}}_{\normalsize\text{\normalsize Recognition model $Q(p|x)$}}$};
    \node[align=center] at ($(s.west)!0.5!(f.east) + (0,-1)$) {$\underbrace{\hspace{5cm}}_{\text{\normalsize Search}}$};
  \end{tikzpicture}
  \caption{Neurally-guided program inference pipeline. Recognition model outputs distribution over program $Q(p|x)$. Program output by enumerative search incorporated into frontier if likelihood $\probability[x|p] > 0$}
  \label{neuralPipeline}
\end{figure}

@@

Mathematically this compression takes the form of
finding the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
We replace this marginal with an AIC approximation
and minimize the following expression,
which can be interpreted as a kind of compression:
\begin{equation}
\underbrace{-\log \probability[\mathcal{D}] + \min_{\theta}\Bigg(-\log
\probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
  X}\underbrace{-\log \sum_{p\in
  \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
\label{AIC}
  \end{equation}
At a high level, our approach is to search locally through the space 
of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
fails to decrease.  These small changes consist of introducing new
candidate $\lambda$-expressions into the DSL.

However, there is a snag with this simple approach:
whenever we add a new expression $e$ to the DSL,
the programs found during waking ($\left\{\mathcal{F}_x \right\}$)
are not written in terms of $e$.
Concretely, imagine we wanted to discover a new DSL procedure for doubling numbers,
after having found the programs
\code{(cons (+ 9 9) nil)} and \code{($\lambda$ (x) (+ (car x) (car x)))}.
As human programmers,
we can look at these pieces of code and recognize that,
if we define a new procedure called \code{double},
defined as \code{($\lambda$ (x) (+ x x))},
then we can rewrite the original programs as
\code{(cons (double 9) nil)} and \code{($\lambda$ (x) (double (car x)))}.
This process is a kind of refactoring
where a new subroutine is defined (\code{double})
and the old programs rewritten
in terms of the new subroutine.
Figure~\ref{mapFactor}A
diagrams this refactoring
process for a more complicated
setting,
where the agent must rediscover the higher-order function \code{map}
starting from the basics of Lisp and the Y-combinator.

We refine our objective to compress
\emph{refactorings} of the programs found during waking, minimizing
\begin{equation}
        -\log \probability[\mathcal{D}] + 
\min_{\theta}\Bigg(-\log \probability[\theta|\mathcal{D}] + \|\theta\|_0 +\sum_{x\in X} - \log \underbrace{\sum_{\substack{p:\\\exists p'\in \mathcal{F}_x: p\manyReduce p'}}}_{\text{Refactors }\mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
\label{factorObjective}
\end{equation}
where $p\manyReduce p'$ is the standard notation for ``expression $p$
evaluates to $p'$ by the rules of $\lambda$-calculus''~\cite{pierce}.
Equation~\ref{factorObjective} captures the idea that we want to add
new components to the DSL while jointly refactoring our old programs
in terms of these new components.  But this joint optimization is
intractable, because there are infinitely many ways of refactoring a
program.  To make refactoring tractable we first limit the degree to
which a piece of code can be refactored: rather than consider every
refactoring, we bound the number of $\lambda$-calculus evaluation
steps separating a refactoring from its original program.  Now the
number of refactoring this is finite but astronomically large: for the
example in Figure~\ref{mapFactor} there are approximately $10^{14}$
possible refactorings -- a quantity that grows exponentially both as a
function of program size and a function of the bound on evaluation
steps. How can we tame this combinatorial explosion?

A structure called a \textbf{version space
  algebra}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}
neatly resolves this exponential growth.  A version space is a
tree-shaped data structure that compactly represents a large set of
programs and supports efficient set operations like union,
intersection, and membership checking.  In
Appendix~\ref{appendixVersion}, we give a dynamic program that takes
as input a program $p$ and then outputs a version space containing the
refactorings of $p$. Figure~\ref{mapFactor}B diagrams a subtree of a
version space containing refactorings of a small program. Our
technique is substantially more efficient than explicitly representing
the space of possible refactorings: for the example in
Figure~\ref{mapFactor}A, we represent the space of refactorings using
a version space with $10^6$ nodes, which encodes $10^{14}$
refactorings.

%% \textbf{Idea 1:} Limit the degree to which
%% a piece of code can be refactored.
%% Instead of considering every refactoring,
%% bound the number of $\lambda$-calculus evaluation steps
%% separating a refactoring from its original program.
%% Formally,
%% we define the set of $n$-step refactorings as:
%% \begin{equation}
%%   R_n(p) = \left\{p'\;:\;p'\underbrace{\reduce p''\reduce\cdots\reduce}_{\text{$\leq n$ times}} p \right\}
%% \end{equation}
%% where $p_1\reduce p_2$ is the standard notation for ``$p_1$ rewrites to $p_2$
%% in one step according to the rules of $\lambda$-calculus''~\cite{pierce}.
%% For example,
%% \begin{align*}
%%   \code{((lambda (x) (x x)) (lambda (y) y))}\reduce&\\
%%   \code{((lambda (y) y) (lambda (y) y))}\reduce&\\
%%   \code{(lambda (y) y)}&
%% \end{align*}
%% Returning to Equation~\ref{factorObjective},
%% this approximation gives the following objective:
%% \begin{equation}
%%         \log \probability[\mathcal{D}] + \argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
%% \label{limitedObjective}
%%   \end{equation}
%% In practice, setting the number of refactoring steps $n$ to 3 suffices
%% to give a competent DSL learning algorithm.  Although the number of
%% refactorings is now finite, it is still prohibitively large, and in
%% fact grows exponentially quickly both as a function of $n$ and as a
%% function of the size of the program being refactored.  For example,
%% for the programs in Figure~\ref{mapFactor}, there are approximately
%% $10^{14}$ possible refactorings.  Next, we show how to tame this
%% exponential explosion.

@@

With this machinery in hand, we now have all the pieces needed to
learn a DSL (Algorithm~\ref{grammarInductionAlgorithm}).
The functions $I\beta(\cdot )$ and \textsc{refactor} construct
a version space from a program and extract the shortest program from a version space, respectively (Algorithm~\ref{grammarInductionAlgorithm}, lines 5-6, 14; see Appendix~\ref{appendixVersion}).
To define the prior distribution over $(\mathcal{D},\theta})$ (Algorithm~\ref{grammarInductionAlgorithm}, lines 7-8), we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
  and $\lambda$ controls how strongly we regularize the size of the DSL.
  We place a symmetric Dirichlet prior over the weight vector $\theta$.

\begin{algorithm}%[H]
  %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
  \caption{DSL Induction Algorithm}
  \label{grammarInductionAlgorithm}
  \begin{algorithmic}[1]
    \State {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
    %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
    \While{true}         
    \State $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta(p)$ \Comment{Construct a version space for each program}
    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
    \State $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
    \EndWhile
  \end{algorithmic}
\end{algorithm}

@@

  \footnotesize
  \begin{scope}[shift={(2,-11)}]  
    \node[draw, rounded corners](u1) at (0,0) {union};
    \node[draw, rounded corners](u2) at ($(0,-1) + (u1.south)$) {\code{(}union\code{ 1)}}; \draw (u1.south) -- (u2.north);
    \node[draw, rounded corners](u21) at ($(-1.3,-1) + (u2.south)$) {\code{(($\lambda$ (x) (x 1)) +)}}; \draw ([xshift=-0.2cm]u2.south) -- (u21.north);
    \node[draw, rounded corners](u22) at ($(2.3,-1) + (u2.south)$) {\code{(($\lambda$ (x) (+ x)) 1)}};  \draw ([xshift=-0.2cm]u2.south) -- (u22.north);

@@

    \node[anchor=north](p) at ($(-6,0) + (u2)$) {\normalsize $\underbrace{\texttt{(+ 1 1)}}_{\text{\normalsize  Program}}$};
    \draw[ultra thick,->] ($(0.5,0) + (p.east)$) -- ($(1.5,0) + (p.east)$);
    \node at ($(1,0.25) + (p.east)$) {Refactors};
    \node(g)[anchor=left] at ($(7,0) + (u12.east)$) {\includegraphics[width = 6cm]{figures/vs.eps}};
    \node(panelB)[ultra thick, rounded corners=0, inner sep=10,outer sep=10, draw, fit=(p) (u1) (vs) (g)] {}; \node at ($(0.5,0) + (panelB.west |- panelB.north)$) {\normalsize\textbf{B}};

@@

Procedural visual concepts are studied across AI and cognitive science
--- Bongard problems~\cite{Moscow}, Raven's progressive
matrices~\cite{raven2003raven}, and Lake et al.'s BPL model of
omniglot~\cite{lake2015human} are prominent examples.  Here we take
inspiration from LOGO Turtle graphics~\cite{turtle}, tasking it with
drawing a modest corpus of images while equipping it with control over
a `pen', along with arithmetic operations on angles and distances.

@@

\subsection{Estimating the continuous parameters $\theta$ of a DSL}\label{mapAppendix}
We use an EM algorithm to estimate the continuous parameters of the DSL, i.e. $\theta$.
Suppressing dependencies on $\mathcal{D}$, the EM updates are
\begin{align}
\label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{q_x}\left[\log \probability\left[p|\theta \right] \right]\\
  q_x(p)&\propto \probability[x|p]\probability[p|\theta]\indicator\left[p\in \mathcal{F}_x \right]
\end{align}
In the M step of EM we will update $\theta$ by instead maximizing a lower bound on $\log \probability[p|\theta]$,
making our approach an instance of Generalized EM.

>>>>

\section{Deep Wake/Sleep Program Induction}\label{overviewSection}

@@

We view these wake sleep/cycles as an approximate inference algorithm that starts with a collection of tasks, written $X$, and infers both a program for each task, as well
as a distribution over programs encoded by a DSL, written
$\mathcal{D}$.  We equip $\mathcal{D}$ with a learned weight vector
$\theta$, and together $(\mathcal{D},\theta)$ define a generative
model over programs (Appendix~\ref{sampleProgram}).  We perform
maximum a posteriori (MAP) inference of the DSL and programs given
$X$.  Writing $J$ for the joint probability of $(\mathcal{D},\theta)$
and $X$, we seek the generative model $\mathcal{D}^*,\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]&\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta &\qquad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
%% where $\probability[x|p]$ scores the likelihood of a task
%% $x\in X$ given a program $p$.\footnote{For example, for list
%%   processing, the likelihood is 1 if the program predicts the observed
%%   outputs on the observed inputs, and 0 otherwise; when learning a
%%   generative model or probabilistic program, the likelihood is the
%%   probability of the program sampling the observation.}
Evaluating Eq.~\ref{intractableObjectives}
entails marginalizing over the infinite set of all programs -- which is impossible.
We make a particle-based approximation to Eq.~\ref{intractableObjectives}
and instead marginalize over a finite \textbf{beam} of programs,
with one beam per task, collectively written $\left\{\mathcal{B}_x \right\}_{x\in X}$.
This particle-based approximation is written $\lowerBound (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$
and acts as a lower bound on the joint:
\begin{align}
 J(\mathcal{D},\theta)\geq \lowerBound  (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\text{, where $|\mathcal{B}_x|$ is small}
\end{align}
Wake and sleep cycles correspond to alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{B}_x \right\}_{x\in X}$  (\textbf{Wake})
and $(\mathcal{D},\theta)$ (\textbf{Consolidation}):
\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the beams.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the beams so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
is large (i.e., programs with high posterior probability).
%Here we use a simple and generic enumerative program synthesis algorithm.
%% which we can accomplish by adding new programs to the beams means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{B}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$.

@@

\noindent\textbf{Sleep (Dreaming): tractably maxing $\lowerBound$ w.r.t. the
  beams.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$Q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
into the beams will most increase $\lowerBound$.

@@

During waking we enumerate programs from the DSL in decreasing order
of their probability according to the recognition model, and then
check if a program $p$ assigns positive probability to a task
($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
$\mathcal{B}_x$. We represent programs as polymorphicly-typed
$\lambda$-calculus expressions, a representation closely resembling
Lisp and functional languages like Haskell and OCaml, including
variables, conditionals, higher-order recursive functions, and the
ability to create new functions.  We `batch' the tasks by randomly
shuffling the training set and playing small minibatches of tasks to
the agent during each wake cycle.

@@

Why enumerate, when the program synthesis community has invented many
sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
We have two reasons:
(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
the space of  programs.
\begin{comment}
  \begin{figure}
    \begin{tikzpicture}[every node/.style={rounded corners,thick}]
      \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
      \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
        e.g., RNN or CNN};
      %    \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
      \node[draw,align = center](s) at ([xshift=2cm]e.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
      \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Beam\\$\mathcal{B}_x$};

      \draw[->] (x.east) -- (e.west);
      \draw[->] (e.east) -- (s.west);
      %    \draw[->] (m.east) -- node(p)[fill=white,align=center,midway,inner sep=0,outer sep=0,rotate=90]{\small $Q_{ijk}(x)$}(s.west);
      \draw[->] (s.east) -- (f.west);
      \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway]{\small likelihood $\probability[x|p]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
      \node[align=center] at ($(0,-1) + (e.center)$) {$\underbrace{\hspace{4cm}}_{\normalsize\text{\normalsize Recognition model $Q(p|x)$}}$};
      \node[align=center] at ($(s.west)!0.5!(f.east) + (0,-1)$) {$\underbrace{\hspace{5cm}}_{\text{\normalsize Search}}$};
    \end{tikzpicture}
    \caption{Neurally-guided program inference pipeline. Recognition model outputs distribution over program $Q(p|x)$. Program output by enumerative search incorporated into beam if likelihood $\probability[x|p] > 0$}
    \label{neuralPipeline}
  \end{figure}
\end{comment}

@@

Mathematically this compression takes the form of
finding the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
We replace this marginal with an AIC approximation~\cite{akaike1998information}
and minimize the following expression,
which can be interpreted as a kind of compression:
\begin{equation}
\underbrace{-\log \probability[\mathcal{D}] + \min_{\theta}\Bigg(-\log
\probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
  X}\underbrace{-\log \sum_{p\in
  \mathcal{B}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
\label{AIC}
  \end{equation}
At a high level, our approach is to search locally through the space 
of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
fails to decrease.  These small changes consist of introducing new
candidate $\lambda$-expressions into the DSL.

However, there is a snag with this simple approach:
whenever we add a new expression $e$ to the DSL,
the programs found during waking ($\left\{\mathcal{B}_x \right\}$)
are not written in terms of $e$.
Concretely, imagine we wanted to discover a new DSL procedure for doubling numbers,
after having found the programs
\code{(cons (+ 9 9) nil)} and \code{($\lambda$ (x) (+ (car x) (car x)))}.
As human programmers,
we can look at these pieces of code and recognize that,
if we define a new procedure called \code{double},
defined as \code{($\lambda$ (x) (+ x x))},
then we can rewrite the original programs as
\code{(cons (double 9) nil)} and \code{($\lambda$ (x) (double (car x)))}.
This process is a kind of refactoring
where a new subroutine is defined (\code{double})
and the old programs rewritten
in terms of the new subroutine.
Figure~\ref{mapFactor}A
diagrams this refactoring
process for a more complicated
setting,
where the agent must rediscover the higher-order function \code{map}
starting from the basics of Lisp and the Y-combinator.

We refine our objective to compress
\emph{refactorings} of the programs found during waking, minimizing
\begin{equation}
        -\log \probability[\mathcal{D}] + 
\min_{\theta}\Bigg(-\log \probability[\theta|\mathcal{D}] + \|\theta\|_0 +\sum_{x\in X} - \log \underbrace{\sum_{\substack{p:\\\exists p'\in \mathcal{B}_x: p\manyReduce p'}}}_{\text{Refactors }\mathcal{B}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
\label{factorObjective}
\end{equation}
where $p\manyReduce p'$ is the standard notation for ``expression $p$
evaluates to $p'$ by the rules of $\lambda$-calculus''~\cite{pierce}.
Equation~\ref{factorObjective} captures the idea that we want to add
new components to the DSL while jointly refactoring our old programs
in terms of these new components.  But this joint optimization is
intractable, because there are infinitely many ways of refactoring a
program.  To make refactoring tractable we first limit the degree to
which a piece of code can be refactored: rather than consider every
refactoring, we bound the number of $\lambda$-calculus evaluation
steps separating a refactoring from its original program.  Now the
number of refactoring this is finite but astronomically large: for the
example in Figure~\ref{mapFactor} there are approximately $10^{14}$
possible refactorings -- a quantity that grows exponentially both as a
function of program size and a function of the bound on evaluation
steps. How can we tame this combinatorial explosion?

A structure called a \textbf{version space
  algebra}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}
neatly resolves this exponential growth.  A version space is a
tree-shaped data structure that compactly represents a large set of
programs and supports efficient set operations like union,
intersection, and membership checking.  In
Appendix~\ref{appendixVersion}, we give a dynamic program that takes
as input a program $p$ and then outputs a version space containing the
refactorings of $p$. Figure~\ref{mapFactor}B diagrams a subtree of a
version space containing refactorings of a small program. Our
technique is substantially more efficient than explicitly representing
the space of possible refactorings: for the example in
Figure~\ref{mapFactor}A, we represent the space of refactorings using
a version space with $10^6$ nodes, which encodes $10^{14}$
refactorings. Appendix~\ref{appendixCompression} specifies how we combine this probabilistic and symbolic machinery to update the DSL.

%% \textbf{Idea 1:} Limit the degree to which
%% a piece of code can be refactored.
%% Instead of considering every refactoring,
%% bound the number of $\lambda$-calculus evaluation steps
%% separating a refactoring from its original program.
%% Formally,
%% we define the set of $n$-step refactorings as:
%% \begin{equation}
%%   R_n(p) = \left\{p'\;:\;p'\underbrace{\reduce p''\reduce\cdots\reduce}_{\text{$\leq n$ times}} p \right\}
%% \end{equation}
%% where $p_1\reduce p_2$ is the standard notation for ``$p_1$ rewrites to $p_2$
%% in one step according to the rules of $\lambda$-calculus''~\cite{pierce}.
%% For example,
%% \begin{align*}
%%   \code{((lambda (x) (x x)) (lambda (y) y))}\reduce&\\
%%   \code{((lambda (y) y) (lambda (y) y))}\reduce&\\
%%   \code{(lambda (y) y)}&
%% \end{align*}
%% Returning to Equation~\ref{factorObjective},
%% this approximation gives the following objective:
%% \begin{equation}
%%         \log \probability[\mathcal{D}] + \argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{B}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
%% \label{limitedObjective}
%%   \end{equation}
%% In practice, setting the number of refactoring steps $n$ to 3 suffices
%% to give a competent DSL learning algorithm.  Although the number of
%% refactorings is now finite, it is still prohibitively large, and in
%% fact grows exponentially quickly both as a function of $n$ and as a
%% function of the size of the program being refactored.  For example,
%% for the programs in Figure~\ref{mapFactor}, there are approximately
%% $10^{14}$ possible refactorings.  Next, we show how to tame this
%% exponential explosion.

@@

@@

  \footnotesize
  \begin{scope}[shift={(-1,-12)}]  
    \node[draw, rounded corners](u1) at (0,0) {union};
    \node[draw, rounded corners](u2) at ($(0,-1) + (u1.south)$) {\code{(}union\code{ 1)}}; \draw (u1.south) -- (u2.north);
    \node[draw, rounded corners](u21) at ($(-1.3,-1) + (u2.south)$) {\code{(($\lambda$ (x) (x 1)) +)}}; \draw ([xshift=-0.2cm]u2.south) -- (u21.north);
    \node[draw, rounded corners](u22) at ($(2.3,-1) + (u2.south)$) {\code{(($\lambda$ (x) (+ x)) 1)}};  \draw ([xshift=-0.2cm]u2.south) -- (u22.north);

@@

    \node[anchor=north](p) at ($(0,2.5) + (u1)$) {\normalsize Program: \texttt{(+ 1 1)}};
    \draw[ultra thick,->] ($(0,-0.1) + (p.south)$) --node[sloped, above, inner sep=5]{Refactors} ($(0,0.1) + (u1.north)$);
    \node(g)[anchor=left] at ($(7,0) + (u12.east)$) {\includegraphics[width = 6cm]{figures/vs.eps}};
    \node(panelB)[ultra thick, rounded corners=0, inner sep=10,outer sep=10, draw, fit=(p) (u1) (vs) (g)] {}; \node at ($(0.5,0) + (panelB.west |- panelB.north)$) {\normalsize\textbf{B}};

@@

Procedural visual concepts are studied across AI and cognitive science
--- Bongard problems~\cite{Moscow}, Raven's progressive
matrices~\cite{raven2003raven}, and Lake et al.'s BPL model of
omniglot~\cite{lake2015human} are prominent examples.  Here we take
inspiration from LOGO Turtle graphics~\cite{turtle}, tasking our agent
with drawing a modest corpus of images while equipping it with control
over a `pen', along with arithmetic operations on angles and
distances.

@@

\subsubsection{Estimating the continuous parameters $\theta$ of a DSL}\label{mapAppendix}
We use an EM algorithm to estimate the continuous parameters of the DSL, i.e. $\theta$.
Suppressing dependencies on $\mathcal{D}$, the EM updates are
\begin{align}
\label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{q_x}\left[\log \probability\left[p|\theta \right] \right]\\
  q_x(p)&\propto \probability[x|p]\probability[p|\theta]\indicator\left[p\in \mathcal{F}_x \right]
\end{align}
In the M step of EM we will update $\theta$ by instead maximizing a lower bound on $\log \probability[p|\theta]$,
making our approach an instance of Generalized EM.
